### 2016-07-25 drafts/writeup_07_25_16.tex
starting writeup

\documentclass{article}

\usepackage[numbers, compress]{natbib}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{amsthm,amsmath,amssymb}
\usepackage{macros}
\usepackage{subcaption}
\usepackage[textfont=small, labelfont=small]{caption}
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.png,.jpg,.eps}

\usepackage{algorithm}
\usepackage{algorithmic}

\usepackage[color=yellow]{todonotes}
\usepackage{booktabs}
\usepackage[inline]{enumitem}
\usepackage{verbatim}

\usepackage[margin=1in]{geometry}

\usepackage{setspace}

\title{Discrete and continuous latent states of neural activity in \textit{Caenorhabditis Elegans}}

\author{Scott W. Linderman 
  and
  Liam Paninski \\
  Columbia University
}

\begin{document}

\singlespacing
\maketitle

\begin{abstract}
  Recent advances in neural recording technologies have enabled
  simultaneous measurements of the majority of head ganglia neurons in
  immobilized C. elegans \cite{kato2015global}. Moreover, since some
  neurons are known to reliably indicate the onset or offset of
  particular behaviors, like ventral and dorsal turns, behavioral
  state can be decoded from the simultaneous population recordings.
  These datasets provide unique visibility into the relationship
  between neural activity and behavior.  While it seems clear that
  activity is inherently lower dimensional than the number of neurons
  due to strong correlations between cells, the nature of the latent brain state
  remains unclear. For example, is brain state better thought of as
  discrete or continuous, or perhaps a combination of the two? Does it
  obey linear or nonlinear dynamics?  We propose a generative approach
  to probing these questions. We model the neural activity as a
  \emph{switching linear dynamical system} (SLDS), with both discrete
  and continuous latent states, and conditionally linear dynamics.  We
  then analyze the posterior distribution over states implied by the
  neural recordings and find that the discrete states correspond to
  stereotypical motor sequences. In contrast to previous work, these
  states are exposed in an entirely unsupervised manner.
\end{abstract}

% Body
\onehalfspacing

\section{Model}

Assume the instantaneous neural activity at time~$t$ for a population
of~$N$ neurons is represented as a vector,~$\by_t \in \reals^N$. In
calcium imaging settings, the entries in this vector may be
instaneous~$\Delta F/F$ measurements, or another signal that captures
neural activity. In this experiment, we use the smoothed time
derivative of~$\Delta F/F$. Over the course of an experiment, we
measure a sequence of vectors, which we combine into a matrix denoted
by~$\by_{1:T}$.

Our model is based on the following assumptions: (i) the instantaneous
neural activity,~$\by_t$, reflects an underlying, low-dimensional
latent state; (ii) this state has a discrete component,~$z_t \in \{1, \ldots, K\}$,
and a continuous component,~$\bx_t \in \reals^D$; (iii) the continuous
latent state has linear dynamics governed by the corresponding discrete
latent state; and (iv) the observed neural activity is a linear function
of the underlying states with additive Gaussian noise.

\TODO{Discuss the motivation for these modeling assumptions.}

These assumptions
are combined in a switching linear dynamical system, which we formalize 
with the following generative model:
\begin{align}
  p(\by_{1:T}, \bx_{1:T}, \bz_{1:T} \given \bTheta) &= 
  \prod_{t=1}^T
  p(z_t \given z_{t-1}, \bTheta) \, 
  p(\bx_t \given z_{t-1}, \bx_{t-1}, \bTheta) \, 
  p(\by_t \given z_t, \bx_t, \bTheta).
\end{align}
Our beliefs about the dynamics of these latent states are encoded in the 
form of the conditional distributions for~$z_t$ and~$\bx_t$. First,
we assume the discrete states follow a Markov process,
\begin{align}
  p(z_t \given z_{t-1}, \bTheta) &\sim \distDiscrete(\bpi^{(z_{t-1})}).
\end{align}
Next, we imbue the continuous latent states with linear Gaussian dynamics,
\begin{align}
  p(\bx_t \given \bx_{t-1}, z_{t-1}, \bTheta) 
  &\sim \distNormal(\bA^{(z_{t-1})} \bx_{t-1} + \bb^{(z_{t-1})}, \bQ^{(z_{t-1})}).
\end{align}
Finally, we impose the assumption of linear observations via the conditional
distribution,
\begin{align}
  p(\by_t \given \bx_{t}, z_{t}, \bTheta) 
  &\sim \distNormal(\bC^{(z_{t})} \bx_{t} + \bd^{(z_{t})}, \bR^{(z_t)}).
\end{align}
Thus, the parameters of the model are,
\begin{align}
  \bTheta &= \left\{ \bA^{(k)}, \bb^{(k)}, \bQ^{(k)}, \bC^{(k)}, \bd^{(k)}, \bR^{(k)}, \bpi^{(k)} \right\}_{k=1}^K .
\end{align}

\TODO{Discuss interpretation of the parameters. Set the stage for visualizing~$\bA^{(k)}$ 
in later sections.}

\section{Inference}
\TODO{Describe block Gibbs updates for~$\bz_{1:T} \given \bx_{1:T},
  \by_{1:T}$ and~$\bx_{1:T} \given \bz_{1:T}, \by_{1:T}$.  Given the
  latent states, the remaining parameters are sampled as in Bayesian
  linear regression. }

\section{Results}

\bibliographystyle{unsrt}
\bibliography{writeup}

\end{document}