### 2017-12-22 drafts/elife/draft.tex
figure captions

\documentclass[11pt]{article}

\input{preamble/preamble.tex}
\input{preamble/preamble_math.tex}
\input{preamble/preamble_acronyms.tex}

\newcommand{\celegans}{\textit{C. elegans}}

% \title{\large Discrete and continuous latent states of neural activity in \textit{Caenorhabditis elegans}}
\title{Hierarchical recurrent models reveal latent states of neural activity in \textit{C. elegans}}
\author{Scott Linderman$^{\text{1}}$,
  Annika Nichols$^{\text{2}}$,
  David Blei$^{\text{1}}$,
  Manuel Zimmer$^{\text{2}}$,
  and
  Liam Paninski$^{\text{1}}$
  \\
  $^{\text{1}}$Columbia University,
  $^{\text{2}}$Research Institute of Molecular Pathology (IMP), Vienna Biocenter
}

\begin{document}

\doublespacing

\maketitle

\begin{abstract}
  Recent advances in neural recording technologies have enabled
  simultaneous measurements of the majority of head ganglia neurons in
  both immobilized and freely-behaving
  \celegans~\citep{schrodel2013brain, prevedel2014simultaneous,
    nguyen2016whole}.  The dynamics of neural activity shed light on
  how \celegans~processes sensory information and generates motor
  activity.  To understand these dynamics, we develop recurrent state
  space models that decompose complex time-series into segments with
  simple, linear dynamics. We incorporate these models into a robust,
  hierarchical framework for combining information across whole-brain
  recordings of many worms.  Using this framework, we reveal latent
  states of population neural activity, along with the discrete
  behavioral modes that drive dynamics in this latent state space.  We
  find stochastic transition patterns between these discrete
  behavioral modes, and we see that transition probabilities are
  determined by a combination of current brain state and environmental
  cues.  In addition to quantifying neural dynamics, this
  probabilistic framework aids in neural identification---currently a
  laborious, manual task---and reveals clusters of neurons that are
  similarly tuned in latent state space.  Finally, we find a
  significant overlap between our inferred states and the
  manually-labeled states of~\citet{kato2015global}
  and~\citet{nichols2017global}, which were shown to correspond to
  different behaviors, like forward crawling, reversals, and
  turns. Our methods automatically discover and quantify these
  behaviorally-meaningful states directly from neural activity,
  yielding powerful new tools for neuro-behavioral analysis.
\end{abstract}

\clearpage

\section*{Introduction}

\clearpage

\section*{Results}

\subsection*{A probabilistic framework that combines information across
  trials to learn canonical dynamics}

We seek a parsimonious characterization of the dynamics of neural
activity in the head ganglia of \celegans. Recent advances in optical
imaging offer simultaneous recordings of most of the head ganglia
neurons, but we must surmount three main challenges in order to learn
a dynamics model from these data.  First, we need to design a model
flexible enough to capture nonlinear dynamics yet interpretable enough
to reveal simplifying structure when it exists.  Moreover, any model
will necessarily be an approximation to the true neural dynamics, and
as such we need to be robust to data that does not quite match model
predictions.  Finally, our model must combine partial recordings from
separate organisms to learn shared neural dynamics yet still allow for
individual variability. We construct a recurrent, robust, and
hierarchical state space model that addresses these challenges,
decomposing complex dynamics into a collection of simpler pieces that
can be learned from noisy, partial recordings with trial-to-trial
variability.

We study the data presented in~\citet{kato2015global} and
\citet{nichols2017global}, which consist of a collection of optical
recordings of calcium fluorescence in immobilized \celegans. The
recordings are 18 minutes long with a frame rate of about 3Hz.  On
average, about 100 single units are extracted from each
recording. \celegans~is genetically stereotyped and each of its 302
neurons has a unique label (e.g. \textsf{AVAL}). On average, 30 single
units could be unambiguously labeled in each
recording. Fig.~\ref{fig:model}a illustrates this type of data. The
three panels represent recordings from three separate worms, and the
rows correspond to different neuron labels.  Neural activity is
measured for only a subset of neurons in each worm.

\begin{figure}[t!]
\centering%
\includegraphics[width=6in]{figures/figure1} 
\caption{
  \textit{A probabilistic framework for combining
  information across multiple worms in order to learn a shared
  dynamical system for \celegans.}
  \textbf{a.}~The data consists of observed neural activity collected
  from many different worms.  In each recording we observe a different
  subset of labeled neurons in the head ganglia.
  \textbf{b.}~We model the observed activity as a linear function of
  an underlying, continuous, and typically low-dimensional latent state,
  here shown as a two-dimensional signal for each recording.
  \textbf{c.}~The dynamics of these continuous states are governed by
  a co-evolving discrete state.
  \textbf{d.}~These discrete states index into a library of linear
  dynamics functions for each worm, here shown as maps from~$\reals^2$ to~$\reals^2$.
  Each dynamics function is selectively used in different regions of
  continuous latent space, as shown by the different intensities of
  the vector field.
  \textbf{e.}~To increase robustness to model misspecification, we
  use a multivariate-t distribution for the noise in the continuous
  state updates.
  \textbf{f.}~Each worm has its own dynamics library, but they are
  constrained to be close to the shared, canonical library.  This
  combines information to learn shared structure while also allowing
  for worm-to-worm variability.
  \textbf{g.} Likewise, we allow each worm to have its own observation
  variances to account for individual differences that may arise from,
  for example, variable levels of fluorescent protein expression.
  Missing neurons are denoted by gray boxes. 
}
\label{fig:model}
\end{figure}

% To model this data, we build on switching linear dynamical systems
% (SLDS) \citep{chang1978state, ackerson1970state, hamilton1990analysis,
%   ghahramani1996switching, murphy1998switching}, a type of state space
% model that views data as a projection of a low-dimensional latent
% state that switches between linear dynamical regimes.  Formally,
% let~$y_t \in \reals^N$ denote the vector of neural activity observed
% at time~$t$.  The expected activity is given by,~$\E[y_t] = g(x_t)$,
% where~$x_t \in \reals^D$ is a continuous latent state
% and~$g: \reals^D \to \reals^N$ is a linear map. We assume~$D \ll N$,
% reflecting the low-dimensional nature of neural activity. SLDS
% simplify the dynamics of the continuous latent states by introducing a
% discrete latent state~$z_t \in \{1, \ldots, K\}$ that indicates the
% current dynamical regime.  Each discrete regime~$k$ is associated with
% a linear map~$f_k: \reals^D \to \reals^D$. The discrete state~$z_t$
% specifies which map to use in order to propagate~$x_t$ to the next
% time step.  Specifically,~$\E[x_{t+1}] = f_{z_t}(x_t)$. By switching
% between different linear dynamical regimes, the SLDS captures highly
% nonlinear dynamics.  We seek to infer the continuous and discrete
% latent states and simultaneously learn the linear maps, the
% distribution of the noise in the observations and the continuous
% states, and the dynamics that govern how the discrete latent states
% evolve.
To model this data, we build on switching linear dynamical systems
(SLDS) \citep{chang1978state, ackerson1970state, hamilton1990analysis,
  ghahramani1996switching, murphy1998switching}, a type of state space
model that views data as a projection of a low-dimensional latent
state that switches between linear dynamical regimes.  The expected
activity is a linear function of an underlying, continuous latent
state, illustrated as two-dimensional time series for each worm in
Fig.~\ref{fig:model}b.  We typically assume this latent state is lower
dimensional than the number of observed neurons. SLDS simplify the
dynamics of the continuous latent states by introducing discrete
latent states that indicate the current dynamical regime
(Fig.~\ref{fig:model}c).  Each discrete regime is associated with a
linear map (Fig.~\ref{fig:model}d), and the discrete state specifies
which map to use in order to propagate the continuous latent state to
the next time step.  By switching between different linear dynamical
regimes, the SLDS captures highly nonlinear dynamics.  We seek to
infer the continuous and discrete latent states and simultaneously
learn the linear maps, the distribution of the noise in the
observations and the continuous states, and the dynamics that govern
how the discrete latent states evolve.

% Recurrent model
Standard SLDS treat the discrete latent states as a simple Markov
process: the next state depends only on the current state. However,
this neglects the more nuanced ways in which discrete and continuous
states may co-evolve. For example, the next discrete state may depend
on both the current discrete state and the current location in
continuous latent space; intuitively, different dynamical regimes may
be employed in different regions of continuous latent space. These
types of dependencies are captured by what are variously called
hybrid~\citep{paoletti2007identification},
augmented~\citep{barber2006expectation}, or \emph{recurrent} state
space models~\citep{linderman2017recurrent}. The maps in
Fig.~\ref{fig:model}d illustrate how these dependencies may modulate
the probability of different dynamics depending on the current
location in continuous latent space. We employ these recurrent models
to learn dynamical regimes that are selectively employed as a
function of current continuous brain state.

% Robust model
While recurrent SLDS can approximate highly nonlinear systems, they
are still misspecified. \emph{Robust} methods~\citep{huber1981robust}
ameliorate this concern by lessening the penalty on observations that
do not quite match the dynamics of the current regime.  We achieve
this robustness by replacing the standard Gaussian noise model with a
multivariate-t distribution~\citep{lange1989robust}.  The heavy tails
of this distribution Fig.~\ref{fig:model}e compares a multivariate-t
distribution to a Gaussian with the same mean; though they have
the same covariance shape, the tails of the robust model fall off
more slowly than those of the Gaussian, allowing for outliers and a
degree of model misspecification.

% Hierarchical model
The third component of our framework is a \emph{hierarchical} model
for learning canonical dynamics from partial recordings collected from
many worms.  Since each recording contains a different subset of
labeled neurons, our first step is to align the data, treating the
activity of neurons that were not found in a given recording as
missing data that must be inferred.  We handle this by integrating
over possible activity of the missing neurons, analytically
marginalizing out this missing data.  The second challenge is
combining information from many recordings while also allowing for
worm-to-worm differences.  For example, the neural activitiy may have
larger variance in some recordings than in others due to different
levels of expression of the fluorescent protein.  We allow for this by
having a separate variance for each worm and neuron, as shown in
Fig.~\ref{fig:model}g.  More importantly, the dynamics of neural
activity may be slightly different from worm to worm.  We account for
this by introducing a canonical library of dynamical regimes
(Fig.~\ref{fig:model}f) that are shared by all worms as well as a
unique library for each individual worm (Fig.~\ref{fig:model}d);
however, we constrain the per-worm dynamics to be close to the
canonical dynamics.

% Inference
We fit the the canonical and per-worm dynamics libraries, the
parameters of the dynamics and observation noise models, and the
discrete and continuous latent states with approximate Bayesian
inference as described in the Supplementary Material. Briefly, we
separate the problem into two steps: learning the continuous latent
states and the mapping from continuous latent states to observations,
then learning the dynamics libraries that govern the continuous latent
states and the correspdonding discrete state sequences.  Both steps are
performed with a block Gibbs sampler, leveraging the structure of the
probabilistic model to perform efficient conditional updates of the
various parameters.

\subsection*{Emission model aids in identification and reveals clusters of neurons}

\begin{figure}[t!]
\centering
\includegraphics[width=6in]{figures/figure2} 
\caption{ \textit{The mapping from continuous latent states to
    observations aids in neural identification and reveals clusters of
    neurons.}
  \textbf{a.}~The observed neural activity (solid black lines) in the
  first worm of the \citet{kato2015global} dataset along with the
  smoothed activity given by the model (blue lines). We use
  correlations learned across the population to predict the activity
  of unobserved neurons (those with dotted black lines).
  \textbf{b.}~These predictions are a useful aid for neural identification.
  We withheld the labels of a fraction of neurons and measured the similarity
  between their activity and the models predictions.  The similarity was
  significantly higher for the correct assignments as compared to incorrect
  assignments.
  \textbf{c.}~The probabilistic model offers a principled means of assessing
  the dimensionality of neural activity, even though only a subset of neurons
  are labeled on any given trial.  We find that ten latent dimensions yields
  the highest test likelihood, and that allowing per-worm observation variances
  yields substantial improvements in test performance.
  \textbf{d.}~The per-worm variances show that observation noise can
  vary substantially from one worm to another. Gray squares denote missing
  neurons.
  \textbf{e.}~We can view the linear mapping from latent states to observations
  as either a matrix or as an embedding of neurons.
  \textbf{f.}~Viewed as an embedding, we see that neurons cluster into
  correlated groups with similar tuning to the continuous latent state.
  \textbf{g.}~These groups roughly correspond to neurons at different
  locations along the worm's body.  }
\label{fig:emissions}
\end{figure}

The parameters of the learned model offer novel insights into the
structure of neural activity in the head ganglia that can only be
obtained by pooling information across many partial recordings.  With
this aggregated information, we can smooth noisy observations and
predict the activity of unobserved neurons. \citet{kato2015global}
used a smoothing procedure to estimate temporal principal components.
By contrast, we fit our model directly to the first-order differences
in calcium fluorescence; our only preprocessing step is to correct for
photo-bleaching. Fig~\ref{fig:emissions}a show our results for the first
of the five worms in this dataset. We see that the model smooths out
noise in the observed activity and, moreover, makes predictions about
the activity of neurons that could not be identified in this worm.

The predicted activity is based on the correlations between neurons
learned from many recordings.  These predictions are a useful aid for
identifying unlabeled neurons.  For example, we see that
\textsf{SMBDL} was not found in the first worm
(Fig.~\ref{fig:emissions}a). If the activity of one of the unlabeled
neurons closely matches the model's predictions for \textsf{SMBDL},
then this provides some evidence in favor of assigning it this
label. We tested this idea by witholding the labels of some
confidently-identified neurons and comparing their activity to the
model's predicted activity for all unlabeled neurons.  We measured the
similarity between the actual and predicted activity and found that
correct label, or ``candidate,'' tends to be much more similar than
other incorrect candidates, as summarized in
Fig.~\ref{fig:emissions}b.  Combined with other side-information, like
the location of the neuron, this provides a powerful aid for identifying
neurons.  The more neurons that we can identify, the better we will be
able to learn about the dynamics of neural activity.

Even though each individual recording contains only a subset of neurons,
the hierarchical model can still estimate the dimensionality of
neural activity in the entire head ganglia.  We reserve the last 20\% of
the data (3.6 min) for testing and estimate the dimensionality
of neural activity based on the marginal log likelihood the model assigns
to this held-out data. We find that the test likelihood peaks with a
ten-dimensional latent state (Fig.~\ref{fig:emissions}c).  Moreover, the likelihood is substantially
higher in the model that includes per-worm observation variances
compared to the model with shared variances for all worms. Fig.~\ref{fig:emissions}d
shows the inferred variance for each worm in the \citet{kato2015global}
dataset.  Neurons that were not identified are marked with a gray box.
This shows the extent of the missing data and the need for hierarchical
modeling.

The linear mapping from latent states to observations reveals clusters
of similarly tuned neurons.  The mapping maybe viewed as either an $N$
(neurons) by $D$ (latent dimensions) matrix or as a collection of~$N$
vectors in $D$-dimensional space, i.e. as an embedding of neurons.
These two views are shown in Fig.~\ref{fig:emissions}d. We note that
many vectors point in similar directions, which implies that there are
groups of correlated neurons.  We clustered the neurons based on the
cosine similarity between their embedding vecors and found five clusters
of neurons shown in Fig.~\ref{fig:emissions}f (order of neurons is the same
as in Fig.~\ref{fig:emissions}a). To make sense of these clusters, we then
looked at their spatial distribution along the head-to-tail axis of the
worm's body.  We found that clusters 2-4 correspond to activity closer to
the head and working backward toward the tail.  The last and largest cluster
contains many neurons in the nerve ring (location 0.15) \todo{check} as well
as some posterior neurons. \todo[inline]{discuss functional significance of clusters.}

\subsection*{SLDS automatically parse data into discrete states with simple dynamics}

\begin{figure}[t!]
\centering
\includegraphics[width=6in]{figures/figure3} 
\caption{ \textit{The hierarchical SLDS decomposes continuous states
    into segments with simple linear dynamics.}
  \textbf{a.}~Three dimensions of the inferred continuous latent states
  for the first worm, color coded by the corresponding discrete latent
  state.
  \textbf{b.}~We select the number of discrete states and justify the
  recurrent and robust extensions by measuring the marginal likelihood
  of held-out test data. We find that the robust, recurrent model
  with eight states performs best.
  \textbf{c.}~The inferred segmentations align with the manual segmentations
  of~\citet{kato2015global}.
  \textbf{d.}~Moreover, the inferred states are in close correspodence with
  the manually labeled states, which were shown to indicate different behaviors.
  \textbf{e.}~Each discrete state is associated with a canonical set of linear
  dynamics, which can be viewed as mappings from~$\reals^10$ to~$\reals^10$.
  We show the mappings for the first three dimensions here, with vectors placed
  at locations where each state was used.  We see a variety of rotational
  states that can be combined to yield dynamics like those in~\textbf{a.}.
  \textbf{f.}~We simulate these linear dynamics from a variety of starting
  locations for the fourth (green) and highlight four for closer inspection.
  \textbf{g.}~These four trajectories give rise to different activations of
  the neural clusters. We see see a dip then rise in cluster 3, which consists
  of neurons related to ventral turns, consistent with the manual labeling
  of the green state in~\textbf{d}.
}
\label{fig:syllables}
\end{figure}

Having learned a mapping from latent state space to observed activity,
we can now reason about the dynamics of these continuous latent
states. Fig.~\ref{fig:syllables}a plots the trajectory of the first
worm's neural activity in the first three dimensions of the continuous
state space, color coded by the discrete regime inferred under the
SLDS.  We see that different regimes correspond to distinct loops
through latent space.  Compare this with Fig. 4b
of~\citet{kato2015global}---they achieved a similar result by manually
segmenting smoothed trajectories in the space spanned by the first
three principal components.  In contrast, our probabilistic framework
is automatic, operates on ten-dimensional latent states, and requires
no additional smoothing.

Moreover, we can use the likelihood of held-out test data to determine
the number of discrete regimes and assess the value of the robust and
recurrent model extensions, as shown in Fig.~\ref{fig:syllables}b.
While the non-robust models with Gaussian dynamics noise achieve
higher training likelihood, the robust models generalize better to
test data.  Moreover, the recurrent models provide further improvement
over the standard Markovian models.  We find that test likelihood is
maximized with eight latent states, which matches the number manually
chosen by~\citet{kato2015global}.

Not only does the SLDS find the same number of states, it also finds a
close correspondence between the manually labeled and inferred states.
Fig.~\ref{fig:syllables}c shows ten-dimensional latent states and the
manual and inferred segmentations for a 5.5 minute section of data.
The manual and inferred states often change at similar times. We also
see instances where there appears to be a change in continuous state
dynamics that are identified by the SLDS but missed in the manual
segmentation (e.g. around $t=700$ seconds).  Fig.~\ref{fig:syllables}d
quantifies the overlap in states for the first worm and for all worms.
Many inferred states are in one-to-one correspondence with manually
labeled states, and the remainder tend to be split between two manual
states.  This confirms that the SLDS is a useful tool for automatically
segmenting brain data.

Beyond simply automating a manual process, the SLDS also provides a
quantitative description of the different states.  In particular, each
state is associated with a set of linear dynamics.
Fig.~\ref{fig:syllables}e shows the canonical dynamics associated with
each of the eight discrete states. The arrows in these vector fields
show where the continuous state is mapped to under the corresponding
discrete state (for the first three latent dimensions).  Arrows are
only drawn at those locations where the discrete state was actually
employed.  We see that these dynamics capture the rotational patterns
of the various loops.

We use these dynamics to simulate how brain activity would evolve
under the different states. Fig.~\ref{fig:syllables}f shows 200
simulations of state 4 from various starting locations.  Four
simulations are highlighted in particular, as denoted by the outlined
circle, triangle, square, and pentagon.  We then map these
trajectories into the space of neural activity and show how they would
activate the five clusters shown in Fig.~\ref{fig:emissions}.  We see
that state 4 corresponds to a pronounced dip followed by a rise in the
activity of cluster 3, alongside a slow ramp in the activity of
clusters 1 and 2. Referring back to Fig~\ref{fig:emissions}a, we see
that cluster 3 consists of \textsf{ALA} and three symmetric pairs of
neurons (\textsf{ASK}, \textsf{RIV}, and \textsf{SMDV}).  The~\textsf{ASK}
pair are sensory neurons responsible for initiating local search behavior,
the \textsf{RIV} pair are polymodal inter- and motor-neurons that specify
the ventral bias of turns, and the \textsf{SMDV} are motor neurons
that initiate ventral turns. As state 4 is in close correspondence with
the ``ventral turn'' state labeled by~\citet{kato2015global}, it makes
sense that these neurons would be activated.

\subsection*{Recurrent models learn spatially localized states and transition boundaries}

\begin{figure}[t!]
\centering
\includegraphics[width=6in]{figures/figure4} 
\caption{ \textit{Discrete state transitions show a non-Markovian,
    chain-like structure.}  \textbf{a.}~Counts of inferred transitions
  from one state to another. The states are permuted to emphasize the
  feed-forward structure.  \textbf{b.}~An alternative view of the
  transitions; each state is a node in a graph and edges denote
  high-probability transitions.  This shows transitions between
  forward and reverse crawling linked by dorsal and ventral turns.
  \textbf{c.}~The standard SLDS assumes Markovian transitions, which
  imply a geometric distribution of state durations (dotted black
  lines), but Fig.~\ref{fig:syllables} clearly demonstrates that state
  transitions occur near the origin.  The recurrent SLDS incorporates
  this dependency and generates state durations (solid black lines)
  that are a much better fit to those inferred from the data (colored
  bars).  }
\label{fig:recurrent}
\end{figure}

The final and most critical component of the dynamical model for
neural activity in the head ganglia is the discrete transition
distribution.  In contrast to standard SLDS, we allow these transition
probabilities to depend on the location in continuous state space (the
\emph{recurrent} dependency) as well as the preceding discrete state
(the \emph{Markovian} dependency).  Fig.~\ref{fig:recurrent}a shows
the contribution of the Markovian term.  Each cell in this matrix
denotes the number of transitions from one state to another. We are
only counting transitions in which the state changes, hence the
diagonal is zero.  We have permuted the states to emphasize the
largely feed-forward structure.

Fig.~\ref{fig:recurrent}b shows the strongest entries in this matrix
in an alternative form.  We have illustrated each state as a node in a
graph, and labeled the states based on their correspondence with the
manually labeled states of~\citet{kato2015global} (see
Fig.~\ref{fig:syllables}d).  This reveals a chain-like transition
structure in which slow and forward (\textsf{FWD}) crawling are linked
to reverse (\textsf{REV1}, \textsf{REV2}) and sustained reverse
(\textsf{REVSUS}) via dorsal turns (\textsf{DT}). From sustained
reverse crawling, the worm may execute either dorsal or ventral turns
(\textsf{VT}).

To accurately model these dynamics, the changes must occur at precise
locations in continuous latent space.  Namely,
Fig.~\ref{fig:syllables}a shows that the loops tend to start and end
at the origin, and we rarely see transitions when the continuous
latent state is far from zero. This manifests in signature duration
distributions for the various states.  Fig.~\ref{fig:recurrent}c plots
these distributions for each of the eight states.  Superimposed on
top, we show the predicted distributions with the recurrent SLDS
(solid black lines) compared with the predictions of the standard,
Markovian SLDS (dotted black lines). The Markovian model knows nothing
of the continuous latent state and implies a geometric duration
distribution with monotonically decreasing probability. By contrast,
the recurrent model learns to transition near the origin, which
gives rise to characteristic duration distributions that closely
match the empirical distributions.

\subsection*{Oxygen level modulates transition probabilities but not dynamics or emissions}

\begin{figure}[t!]
\centering
\includegraphics[width=6in]{figures/figure5} 
\caption{\textit{Incorporating exogenous factors like oxygen level
    into the dynamics model yield differential responses in worms from
    different genetic strains and developmental stages.}
  \textbf{a.}~The observed neurons and their inferred observation variances
  for each of the 41 worms in the~\citet{nichols2017global} dataset.
  \textbf{b.}~Worms of different genetic strains (N2, npr1) and developmental
  stages (pre-lethargus, lethargus) exhibit different levels of overall
  state usage, with blue and yellow much more likely in the lethargus stage.
  \textbf{c.}~As above, the inferred states closely match the manually
  labeled states of~\citet{nichols2017global}, with blue and yellow mapping
  onto the forward (\textsf{Fwd}) and quiescent (\textsf{Q}) states.
  \textbf{d.}~We extended the SLDS to allow exogenous factors like oxygen
  level influence discrete transition probabilities.
  \textbf{e., f.}~Upon fitting the model, we find that oxygen has a limited
  effect on N2 worms and a much stronger effect on npr1 worms. In npr1 worms,
  increasing the oxygen level decreases the probability of quiescent and
  forward states and elicits reversals and turns.
}
\label{fig:o2}
\end{figure}

The preceding results demonstrated the efficacy of hierarchical,
recurrent, and robust SLDS for learning interpretable dynamics of
neural activity from the noisy, partial recordings
of~\citet{kato2015global}.  We now turn to the dataset studied
in~\citet{nichols2017global}, which contains many more worms with
roughly the same number of neurons per recording but is complicated by
additional factors.  First, there are two strains of worms (N2 and
npr-1) at two different stages of development (pre-lethargus and
lethargus); second, the oxygen level is manipulated over the course of
the recordings to elicit different responses.  Here we show how our
framework can be extended to study the effects of these manipulations.

Fig.~\ref{fig:o2}a shows which neurons were labeled in each of the
recordings along with the inferred observation variance. There are 41
worms, roughly ten from each strain and developmental stage
combination (N2 + pre-lethargus, N2 + lethargus, npr1 + pre-lethargus,
and npr1 + lethargus). Some neurons are consistently identified in
almost all worms; others are identified in only a handful.  As before,
our hierarchical framework combines all this information.

Upon fitting our model, we immediately see a difference in the overall
usage of the eight states (Fig.~\ref{fig:o2}b). The worms in the
lethargus stage predominantly use the blue and yellow states.  Again,
we found a close correspondence between our inferred states and the
manually labeled states of~\citet{nichols2017global}
(Fig.~\ref{fig:o2}c). Here, blue and yellow both map strongly onto the
quiescent (\textsf{Q}) state. This is consistent with the
characterization of the lethargus stage---at this stage, worms exhibit
much less crawling behavior. The pre-lethargus worms execute a broader
variety of behaviors, including reversals (\textsf{Rev.}) and dorsal
and ventral turns  (\textsf{DT} and \textsf{VT}, respectively).

A key finding of~\citet{nichols2017global} was that increasing the
oxygen level can elicit activity in lethargus npr1 worms but not in
lethargus N2 worms. We can quantify this effect with our models by
including oxygen level as an exogenous input.  Fig.~\ref{fig:o2}d
shows how the graphical model of the recurrent SLDS changes: now the
discrete state at time~$t$ (red circle) depends on the preceding discrete
state, the preceding continuous state, and the current oxygen level.
\todo[inline]{consider alternative models in which oxygen level
  influences continuous states and observations?}

Fig.~\ref{fig:o2}e and~\ref{fig:o2}f show the inferred effect of
oxygen level for each of the four strain/developmental stage pairs.
As predicted, only the npr1 worms show a much larger response to
oxygen. Increasing the oxygen level increases the probability
of reversals and turning related states (red, orange, purple) and
decreases the probability of forward and quiescent (blue, yellow)
states. Somewhat surprisingly, the magnitude of the effect is
comparable for both pre-lethargus and lethargus npr1 worms, even
though the baseline level of reversals and turns is much higher in
the pre-lethargus worms.

\clearpage

\section*{Discussion}

\clearpage

\bibliography{refs}
\bibliographystyle{abbrvnat}

\clearpage

\appendix

\section{Data}
\label{sec:data}

The data comes in the form of a collection of matrices~$\{\widetilde{Y}^{(w)}\}_{w=1}^W$
for each of~$W$ worms.  The matrix~${\widetilde{Y}^{(w)} \in \reals^{T_w \times N_w}}$
represents the calcium activity of the~$w$-th worm, which consists of~$T_w$
time frames and~$N_w$ neurons.\footnote{Technically, this calcium activity is
a matrix of first-order temporal differences of a
bleaching-corrected~$\Delta F /F$ signal. Unlike~\citet{kato2015global}, we
do not smooth these derivatives. We work with the raw first-order
temporal differences.}  \celegans~is special in that each neuron has a unique
name.  Assigning names to neurons in calcium imaging data is a challenging
task, but typically, out of the~${\sim 100}$ neurons observed in any worm,
we can label about~$20-30$ with high certainty.  As such, we choose to
represent each matrix~$\widetilde{Y}^{(w)}$ in \emph{canonical form} as a
tuple~$(Y^{(w)}, M^{(w)})$, where~${Y^{(w)} \in \reals^{T_w \times N}}$ is
a matrix where each column corresponds to one of the~$N$ neurons that is
identified in at least one of the~$W$ worms, and~${M^{(w)} \in \{0,1\}^{T_w \times N}}$
is a corresponding \emph{mask} matrix that indicates which neurons were
observed in worm~$w$.  If~$M_{:,n}^{(w)} = 1$, neuron~$n$ was labeled in
worm~$w$ and~$Y_{:,n}^{(w)}$ was its observed activity.  If~$M_{:,n}^{(w)}=0$,
this neuron was not observed in worm~$w$, and the corresponding activity
is missing.\footnote{The mask is defined in such a way that neurons can
  be missing for only subsets of time frames, but in our data neurons are
  either labeled or missing.}
% todo: note that we could also (and have also) model the activity of
% unlabeled neurons.  For the final analyses, I'm not doing this though.

\section{Hierarchical Generative Model}

We model the neural activity with a switching linear
dynamical system (SLDS) with three new extensions: (i)~a \emph{hierarchical} model
to share parameters across worms while also allowing for worm-to-worm
variability; (ii)~a \emph{robust} model for dynamics noise to help address
model misspecification; and (iii)~a \emph{recurrent} model to capture
how continuous latent states influence discrete state transition probabilities.
We will introduce the SLDS first and then present each of these extensions
in turn.

\subsection{Switching linear dynamical systems}
\label{sec:slds}
Switching linear dynamical system models (SLDS) break down complex, nonlinear
time series data into sequences of simpler, reused dynamical modes.
By fitting an SLDS to data, we not only learn a flexible nonlinear generative
model, but also learn to parse data sequences into coherent discrete units.

The generative model is as follows. At each time ${t=1,2,\ldots,T}$,
for each worm~$w$,
there is a discrete latent state ${z_t^{(w)} \in \{1,2, \ldots,K\}}$ that
follows Markovian dynamics,
\begin{equation}
  z_{t+1}^{(w)} \given z_t^{(w)}, \{\pi_k\}_{k=1}^K  \sim \pi_{z_t^{(w)}}
  \label{eq:markov_z}
\end{equation}
where ${\{\pi_k\}_{k=1}^K}$ is the Markov transition matrix and
${\pi_k \in [0,1]^K}$ is its $k$th row. 
In addition, a continuous latent state ${x_t^{(w)} \in \reals^D}$ follows
conditionally linear (or affine) dynamics, where the discrete state $z_t^{(w)}$
determines the linear dynamical system used at time $t$:
\begin{align}
  x_{t+1}^{(w)} &= A_{z_{t+1}^{(w)}} x_{t}^{(w)} + b_{z_{t+1}^{(w)}} +  u_t^{(w)},
  &
  u_t^{(w)} &\iid\sim \cN(0,Q_{z_{t+1}^{(w)}}),
  \label{eq:slds_start}
\end{align}
for matrices ${A_k, Q_k \in \reals^{D \times D}}$ and vectors~${b_k\in
\reals^D}$ for~${k=1,2,\ldots,K}$.
Finally, at each time $t$ a linear Gaussian observation $y_t^{(w)} \in \reals^N$
(some entries of~$y_t^{(w)}$ are masked off; recall Section~\ref{sec:data}) is
generated from the corresponding latent continuous state,
\begin{align}
  y_t^{(w)} &= C x_t^{(w)} + d + v_t^{(w)}, & v_t^{(w)} &\iid\sim \cN(0,S),
    \label{eq:slds_end}
\end{align}
for $C \in \reals^{N \times D}$, $S \in \reals^{N \times N}$,
and~$d \in \reals^N$. We denote the rows of~$C$ by vectors~$c_n$.
For simplicity, we assume~$C$,~$d$, and~$S$ to
be shared among all discrete states in our model.  Moreover, we assume
the observation noise is diagonal,
\begin{align*}
  S &= \diag\left(\left[s_{1}, \ldots, s_{N} \right] \right). 
\end{align*}
The system parameters comprise the discrete Markov transition matrix, the
library of linear dynamical system matrices, and the neuron-specific
emission parameters, which we write as
\begin{equation*}
  \theta = \{(\pi_k, A_k, Q_k, b_k)\}_{k=1}^K \cup \{c_n, d_n, s_n\}_{n=1}^N.
\end{equation*}

To learn an SLDS using Bayesian inference, we place conjugate Dirichlet priors
on each row of the transition matrix and conjugate matrix normal
inverse Wishart (MNIW) priors on the linear dynamical system parameters,
Gaussian priors on the rows of the emission matrix,
and inverse gamma priors on the emission noise.
We write this as,
\begin{align*}
  \pi_k &\given \alpha \iid\sim \distDirichlet(\alpha),
  &
  [A_k, b_k], Q_k &\given \lambda \iid\sim \distMNIW(\lambda),
  \\
  [c_n, d_n] &\given \eta \iid\sim \distNormal(\eta),
  &
  s_{n} &\given \eta \iid\sim \mathrm{IG}(\eta).
\end{align*}
where~$[\cdot, \cdot]$ denotes column concatenation, and $\alpha$, $\lambda$,
and~$\eta$ denote appropriate hyperparameters of the transitions, dynamics, and
emissions, respectively.

\end{document}